---
format:
  pdf:
    toc: false
    number-sections: false

fontsize: 10pt

execute:
  echo: true
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(broom)

# Load data
qualtrics <- read_csv(
  "groupproject_data/data/PRCAQualtricsExport_FileC.csv",
  show_col_types = FALSE
)

prolific_a <- read_csv(
  "groupproject_data/data/PRCAProlificExport_FileA.csv",
  show_col_types = FALSE
)

prolific_b <- read_csv(
  "groupproject_data/data/PRCAProlificExport_FileB.csv",
  show_col_types = FALSE
)

# Combine Prolific files
prolific <- bind_rows(prolific_a, prolific_b)
```

```{r}
#| label: score-prca
#| include: false

# Convert PRCA Likert responses to numeric scores
score_likert <- function(x) {
  case_when(
    x == "Strongly disagree" ~ 1,
    x == "Somewhat disagree" ~ 2,
    x == "Neither agree nor disagree" ~ 3,
    x == "Somewhat agree" ~ 4,
    x == "Strongly agree" ~ 5,
    TRUE ~ NA_real_
  )
}

prca_data <- qualtrics |>
  mutate(
    across(
      c(
        Q1, Q2, Q3, Q4, Q5, Q6,
        Q13, Q14, Q15, Q16, Q17, Q18
      ),
      score_likert
    )
  ) |>
  mutate(
    # Group discussion subscale
    prca_group =
      18 - (Q2 + Q4 + Q6) +
      (Q1 + Q3 + Q5),

    # Interpersonal communication subscale
    prca_interpersonal =
      18 - (Q14 + Q16 + Q17) +
      (Q13 + Q15 + Q18),

    # Combined communication apprehension score
    ca_score =
      prca_group + prca_interpersonal
  )
```

```{r}
#| label: merge-age
#| include: false

# Select participant age from Prolific
prolific_age <- prolific |>
  select(
    prolific_id = `Participant id`,
    age = Age
  )

# Merge age with communication apprehension scores
age_data <- prca_data |>
  left_join(
    prolific_age,
    by = c("Q0" = "prolific_id")
  ) |>
  filter(
    !is.na(age),
    !is.na(ca_score)
  )
```

```{r}
#| label: statistical-analysis
#| include: false

# Pearson correlation
age_correlation <- cor.test(
  age_data$age,
  age_data$ca_score,
  method = "pearson"
)

# Ordinary least squares regression
age_model <- lm(
  ca_score ~ age,
  data = age_data
)

# Store results for use in memo text
age_results <- tidy(
  age_model,
  conf.int = TRUE
)

model_results <- glance(age_model)

age_coefficient <- age_results |>
  filter(term == "age")

n_participants <- nrow(age_data)

r_value <- unname(age_correlation$estimate)
age_estimate <- age_coefficient$estimate
age_se <- age_coefficient$std.error
age_t <- age_coefficient$statistic
age_p <- age_coefficient$p.value
age_r2 <- model_results$r.squared
```

**Researcher:** Vanessa Kountakis

## Research Question: Does age predict communication apprehension?


```{r}
#| label: fig-age-ca
#| fig-cap: "Relationship between age and communication apprehension."
#| fig-width: 7
#| fig-height: 4.5
#| fig-align: center

ggplot(
  age_data,
  aes(x = age, y = ca_score)
) +
  geom_point(alpha = 0.5) +
  geom_smooth(
    method = "lm",
    se = TRUE
  ) +
  labs(
    x = "Age",
    y = "Communication Apprehension Score"
  ) +
  theme_minimal()
```

## Answer, Response, + Summary of Results

This analysis examined whether age predicts communication apprehension among `r n_participants` participants using combined group discussion and interpersonal PRCA scores. The scatterplot showed a slight downward trend, but scores varied widely across ages. A Pearson correlation and linear regression found a weak negative relationship, *r* = `r sprintf("%.3f", r_value)`, *b* = `r sprintf("%.3f", age_estimate)`, *p* = `r sprintf("%.3f", age_p)`. Age explained only `r sprintf("%.1f", age_r2 * 100)`% of the variation, so it was not a significant predictor of communication apprehension.


## What Questions or Uncertainties Remain?

A Future analyses could examine whether factors such as transportation habits, employment, education, or other demographic and behavioral characteristics are associated with communication apprehension.