---
format:
  pdf:
    toc: false
    number-sections: false

fontsize: 10pt

execute:
  echo: true
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(broom)
library(knitr)

qualtrics <- read_csv(
  "data/PRCAQualtricsExport_FileC.csv",
  show_col_types = FALSE
)
```

```{r}
#| label: score-prca
#| include: false

score_likert <- function(x) {
  case_when(
    x == "Strongly disagree" ~ 1,
    x == "Somewhat disagree" ~ 2,
    x == "Neither agree nor disagree" ~ 3,
    x == "Somewhat agree" ~ 4,
    x == "Strongly agree" ~ 5,
    TRUE ~ NA_real_
  )
}

prca_data <- qualtrics |>
  mutate(
    across(
      c(
        Q1, Q2, Q3, Q4, Q5, Q6,
        Q13, Q14, Q15, Q16, Q17, Q18
      ),
      score_likert
    )
  ) |>
  mutate(
    prca_group =
      18 - (Q2 + Q4 + Q6) +
      (Q1 + Q3 + Q5),

    prca_interpersonal =
      18 - (Q14 + Q16 + Q17) +
      (Q13 + Q15 + Q18),

    ca_score =
      prca_group + prca_interpersonal
  )
```

```{r}
#| label: transportation-data
#| include: false

transport_levels <- c(
  "Never",
  "0-1 days a month",
  "2-4 days a month",
  "4-8 days a month",
  "8 or more days a month"
)

transport_data <- prca_data |>
  mutate(
    public_transit = factor(
      Q26,
      levels = transport_levels
    ),
    rideshare = factor(
      Q28,
      levels = transport_levels
    ),

    # Numeric coding used only for supplemental Pearson correlations
    public_transit_num = as.numeric(public_transit) - 1,
    rideshare_num = as.numeric(rideshare) - 1
  ) |>
  filter(
    !is.na(ca_score),
    !is.na(public_transit),
    !is.na(rideshare)
  )

n_transport <- nrow(transport_data)
```

```{r}
#| label: anova-analysis
#| include: false

# One-way ANOVA: public transportation
public_anova <- aov(
  ca_score ~ public_transit,
  data = transport_data
)

# One-way ANOVA: rideshare
rideshare_anova <- aov(
  ca_score ~ rideshare,
  data = transport_data
)

# Clean ANOVA results
public_results <- tidy(public_anova)
rideshare_results <- tidy(rideshare_anova)

# Store results for memo text
public_f <- public_results |>
  filter(term == "public_transit") |>
  pull(statistic)

public_p <- public_results |>
  filter(term == "public_transit") |>
  pull(p.value)

rideshare_f <- rideshare_results |>
  filter(term == "rideshare") |>
  pull(statistic)

rideshare_p <- rideshare_results |>
  filter(term == "rideshare") |>
  pull(p.value)

# Tukey post-hoc tests
public_tukey <- TukeyHSD(public_anova)
rideshare_tukey <- TukeyHSD(rideshare_anova)

# Combined ANOVA table for appendix
anova_table <- bind_rows(
  public_results |>
    mutate(Analysis = "Public Transportation"),
  rideshare_results |>
    mutate(Analysis = "Rideshare")
) |>
  mutate(
    Source = case_when(
      term == "Residuals" ~ "Residuals",
      TRUE ~ "Between Groups"
    )
  ) |>
  select(
    Analysis,
    Source,
    df,
    sumsq,
    meansq,
    statistic,
    p.value
  )
```

```{r}
#| label: pearson-analysis
#| include: false

# Supplemental Pearson correlations
public_pearson <- cor.test(
  transport_data$public_transit_num,
  transport_data$ca_score,
  method = "pearson"
)

rideshare_pearson <- cor.test(
  transport_data$rideshare_num,
  transport_data$ca_score,
  method = "pearson"
)

public_r <- unname(public_pearson$estimate)
public_cor_p <- public_pearson$p.value

rideshare_r <- unname(rideshare_pearson$estimate)
rideshare_cor_p <- rideshare_pearson$p.value

correlation_table <- tibble(
  Transportation = c(
    "Public Transportation",
    "Rideshare"
  ),
  r = c(
    public_r,
    rideshare_r
  ),
  p = c(
    public_cor_p,
    rideshare_cor_p
  )
)
```

```{r}
#| label: transportation-plot-data
#| include: false

transport_plot <- transport_data |>
  select(
    ca_score,
    public_transit,
    rideshare
  ) |>
  pivot_longer(
    cols = c(public_transit, rideshare),
    names_to = "transport_type",
    values_to = "frequency"
  ) |>
  mutate(
    transport_type = dplyr::recode(
      transport_type,
      public_transit = "Public Transportation",
      rideshare = "Rideshare"
    ),
    frequency = dplyr::recode(
      as.character(frequency),
      "Never" = "Never",
      "0-1 days a month" = "0–1 days/month",
      "2-4 days a month" = "2–4 days/month",
      "4-8 days a month" = "4–8 days/month",
      "8 or more days a month" = "8+ days/month"
    ),
    frequency = factor(
      frequency,
      levels = c(
        "Never",
        "0–1 days/month",
        "2–4 days/month",
        "4–8 days/month",
        "8+ days/month"
      )
    )
  )
```

**Researcher:** Vanessa Kountakis

## Research Question: Do Transportation Habits Predict Communication Apprehension?

**Answer, Response, + Summary of Results**

A one-way ANOVA found no significant differences in communication apprehension across public transportation frequency groups, *F*(4, 255) = `r sprintf("%.2f", public_f)`, *p* = `r sprintf("%.3f", public_p)`. In contrast, communication apprehension differed significantly across rideshare frequency groups, *F*(4, 255) = `r sprintf("%.2f", rideshare_f)`, *p* < .001. Tukey comparisons showed that participants who used rideshare 2–4 days per month, 4–8 days per month, or 8 or more days per month had significantly lower communication apprehension than participants who never used rideshare.

Supplemental Pearson correlations showed a small negative relationship between public transportation frequency and communication apprehension, *r* = `r sprintf("%.3f", public_r)`, *p* = `r sprintf("%.3f", public_cor_p)`, and a stronger negative relationship between rideshare frequency and communication apprehension, *r* = `r sprintf("%.3f", rideshare_r)`, *p* < .001.

Overall, rideshare frequency showed the clearest relationship with communication apprehension, with more frequent rideshare users generally reporting lower communication apprehension.

**What Questions or Uncertainties Remain?**

Would rideshare frequency remain associated with communication apprehension after accounting for factors such as age, car access, and employment status?

\newpage

# Appendix

## Figure 1: Transportation Habits and Communication Apprehension

```{r}
#| label: fig-transport-ca
#| fig-cap: "Communication apprehension scores across public transportation and rideshare frequency groups."
#| fig-width: 6.5
#| fig-height: 3.2
#| fig-align: center

ggplot(
  transport_plot,
  aes(
    x = frequency,
    y = ca_score
  )
) +
  geom_boxplot() +
  facet_wrap(
    ~ transport_type,
    ncol = 2
  ) +
  labs(
    x = "Frequency of Use",
    y = "Communication Apprehension Score"
  ) +
  theme_minimal(base_size = 9) +
  theme(
    axis.text.x = element_text(
      angle = 25,
      hjust = 1
    )
  )
```

## Table 1: ANOVA Results

```{r}
#| label: tbl-anova

anova_table |>
  mutate(
    sumsq = round(sumsq, 2),
    meansq = round(meansq, 2),
    statistic = round(statistic, 2),
    p.value = case_when(
      is.na(p.value) ~ "",
      p.value < .001 ~ "< .001",
      TRUE ~ sprintf("%.3f", p.value)
    )
  ) |>
  kable(
    col.names = c(
      "Analysis",
      "Source",
      "df",
      "Sum of Squares",
      "Mean Square",
      "F",
      "p"
    ),
    caption = "One-way ANOVA results for transportation frequency and communication apprehension."
  )
```

## Table 2: Supplemental Pearson Correlations

```{r}
#| label: tbl-pearson

correlation_table |>
  mutate(
    r = round(r, 3),
    p = case_when(
      p < .001 ~ "< .001",
      TRUE ~ sprintf("%.3f", p)
    )
  ) |>
  kable(
    col.names = c(
      "Transportation Measure",
      "r",
      "p"
    ),
    caption = "Supplemental Pearson correlations between transportation frequency and communication apprehension."
  )
```

*Note.* Transportation frequency was coded from 0 = Never to 4 = 8 or more days per month for the supplemental Pearson correlations.