---
title: "Com_Apprehension_Transportation"
format: html
editor: visual
---

# Preparation

```{r}

library(tidyverse)
options(conflicts.policy = "depends.ok")
path_data <- "groupproject_sibling_data/cleandata"
```

```{r}
data <- read.csv(here::here(path_data, "minimal_clean.csv")) |> 
  glimpse()
```

```{r}
data_all <- data |>
  mutate(
    Age = as.integer(Age),
    Sex = factor(Sex),
    public_transport = factor(
      public_transport,
      levels = c(0, 1)
    ),
    public_transport_monthly = ordered(
      public_transport_monthly,
      levels = c(0, 1, 2, 3, 4)
    ),
    public_transport_daily = ordered(
      public_transport_daily,
      levels = c(1, 2, 3, 4)
    ),
    ride_share = factor(
      ride_share,
      levels = c(0, 1)
    ),
    ride_share_monthly = ordered(
      ride_share_monthly,
      levels = c(0, 1, 2, 3, 4)
    ),
    ride_share_daily = ordered(
      ride_share_daily,
      levels = c(1, 2, 3, 4)
    )
  ) |>
  glimpse()
```

```{r}
data_all <- data_all |>
  mutate(
    transport_group = case_when(
      ride_share == 0 & public_transport == 0 ~ "Neither",
      ride_share == 1 & public_transport == 0 ~ "Ride share only",
      ride_share == 0 & public_transport == 1 ~ "Public transport only",
      ride_share == 1 & public_transport == 1 ~ "Both"
    ),
    transport_group = factor(
      transport_group,
      levels = c(
        "Neither",
        "Ride share only",
        "Public transport only",
        "Both"
      )
    )
  )
```

# Analysis

### Ride-share users vs non-users

```{r}
ride_share_model <- lm(
  communication_apprehension ~ ride_share,
  data = data_all
)

summary(ride_share_model)
```

Ride-share users had communication apprehension scores that were, on average, 0.62 points lower (per question) than those of non-users, *b* = −0.62, *SE* = 0.13, *t*(250) = −4.96, *p* \< .001.

### Public transport users vs non-users

```{r}
public_transport_model <- lm(
  communication_apprehension ~ public_transport,
  data = data_all
)

summary(public_transport_model)
```

Public transportation users had communication apprehension scores that were, on average, 0.53 points lower (per question) than those of non-users, *b* = −0.53, *SE* = 0.14, *t*(250) = −3.71, *p* \< .001.

### Both (there is overlap)

Some participants use both public transportation and ride-shares, so this model doesn't directly answer our question.

```{r}
transport_model <- lm(
  communication_apprehension ~ ride_share + public_transport,
  data = data_all
)

summary(transport_model)
```

Ride-share users who did not use public transportation had estimated communication apprehension scores 0.51 points lower, on average, than participants who used neither mode, *b* = −0.51, *SE* = 0.14, *t*(249) = −3.62, *p* \< .001.

Public-transport users who did not use ride share had estimated communication apprehension scores 0.26 points lower, on average, than participants who used neither mode, *b* = −0.26, *SE* = 0.16, *t*(249) = −1.66, *p* = .098.

|        Transportation         | Avg. Communication Apprehension/Question |
|:-----------------------------:|:----------------------------------------:|
|            Neither            |                   1.98                   |
|     Public Transportation     |                   1.72                   |
|          Ride Share           |                   1.47                   |
| Public Transport & Ride Share |                   1.21                   |

: \*Public Transportation value uses ALL Ps who use public transport (including those that also use ride-share). Likewise, Ride Share Transportation value uses ALL Ps who use ride-share (including those that also use public transportation).

### Both, using groups to isolate

```{r}
transport_group_model <- lm(
  communication_apprehension ~ transport_group,
  data = data_all
)

summary(transport_group_model)
```

|        Transportation         | Avg. Communication Apprehension/Question |
|:-----------------------------:|:----------------------------------------:|
|            Neither            |                   2.11                   |
|  Public Transportation Only   |                   1.59                   |
|        Ride Share Only        |                   1.17                   |
| Public Transport & Ride Share |                   1.23                   |

**Ride share only vs. neither:**

Participants who used ride share but not public transportation reported communication apprehension scores that were, on average, **0.94 points lower** than participants who used neither mode, *b* = −0.94, *SE* = 0.27, *t*(248) = −3.44, *p* \< .001.

**Public transport only vs. neither:**

Participants who used public transportation but not ride share reported communication apprehension scores that were, on average, **0.52 points lower** than participants who used neither mode, *b* = −0.52, *SE* = 0.21, *t*(248) = −2.46, *p* = .015.

**Both vs. neither:**

Participants who used both ride share and public transportation reported communication apprehension scores that were, on average, **0.88 points lower** than participants who used neither mode, *b* = −0.88, *SE* = 0.16, *t*(248) = −5.33, *p* \< .001.

**Overall model is significant:**

Transportation-use group was significantly associated with communication apprehension, *F*(3, 248) = 10.38, *p* \< .001, Rsqu = .112, so \~11.2% of the variance in communication apprehension is accounted for by which transportation group one falls into.

```{r}
data_all <- data_all |>
  mutate(
    transport_group = relevel(
      transport_group,
      ref = "Public transport only"
    )
  )

model_group <- lm(
  communication_apprehension ~ transport_group,
  data = data_all
)

summary(model_group)
```

Using public transport only as our reference group, we now have a test statistic that directly compares public transport only to ride share only.

Participants who use ride share but not public transportation had communication apprehension scores that were, on average, 0.43 points lower than participants who use public transportation but not ride share; however, this difference was not statistically significant, *b* = −0.43, *SE* = 0.27, *t*(248) = −1.55, *p* = .122.

**Maybe more interesting question: Do Ps who use ride-share, public transport, or both differ in communication apprehension from people who use neither?**

```{r}
data_all <- data_all |>
  mutate(
    any_transport = case_when(
      ride_share == 0 & public_transport == 0 ~ "Neither",
      ride_share == 1 | public_transport == 1 ~ "Any"
    ),
    any_transport = factor(
      any_transport,
      levels = c("Neither", "Any")
    )
  )

model_any <- lm(
  communication_apprehension ~ any_transport,
  data = data_all
)

summary(model_any)
```

Participants who used either ride share, public transportation, or both reported communication apprehension scores that were, on average, 0.82 points lower than participants who used neither mode, *b* = −0.82, *SE* = 0.16, *t*(250) = −5.08, *p* \< .001.

### Analysis including frequency of use.

Probably going to only use monthly since the daily question didn't give participants the option to choose 0 even if they reported never using that mode of transport.

#### Quick EDA

```{r}
data_all |>
  group_by(ride_share_monthly) |>
  summarize(
    n = n(),
    mean_CA = mean(communication_apprehension, na.rm = TRUE),
    sd_CA = sd(communication_apprehension, na.rm = TRUE)
  )



data_all |>
  group_by(public_transport_monthly) |>
  summarize(
    n = n(),
    mean_CA = mean(communication_apprehension, na.rm = TRUE),
    sd_CA = sd(communication_apprehension, na.rm = TRUE)
  )
```

#### Analysis

```{r}
model_ride_frequency <- lm(
  communication_apprehension ~ factor(ride_share_monthly),
  data = data_all
)

model_public_frequency <- lm(
  communication_apprehension ~ factor(public_transport_monthly),
  data = data_all
)

summary(model_ride_frequency)
summary(model_public_frequency)

anova(model_ride_frequency)
anova(model_public_frequency)
```

**Overall model**

Communication apprehension differed significantly across levels of monthly ride-share use, *F*(4, 247) = 9.31, *p* \< .001, Rsqu = .131

```{r}
model_frequency_both <- lm(
  communication_apprehension ~
    factor(ride_share_monthly) +
    factor(public_transport_monthly),
  data = data_all
)

summary(model_frequency_both)
anova(model_frequency_both)
```

There was a significant negative linear trend in communication apprehension across increasing categories of ride-share frequency, after controlling for public-transport frequency, *b* = −0.64, *SE* = 0.19, *t*(243) = −3.47, *p* \< .001.

i.e. **Ps tended to have lower communication apprehension as ride-share use increased, even after accounting for differences in public transportation use.**

AND

There was no evidence of a linear trend in communication apprehension across increasing categories of public-transport frequency after controlling for ride-share frequency, *b* = −0.13, *SE* = 0.15, *t*(243) = −0.86, *p* = .390.

i.e. **After accounting for ride-share frequency, public transportation frequency was not significantly associated with communication apprehension.**

#### Analysis treating frequency categories as numeric rather than ordinal

```{r}
data_all <- data_all |>
  mutate(
    ride_frequency = as.numeric(as.character(ride_share_monthly)),
    public_frequency = as.numeric(as.character(public_transport_monthly))
  )

model_frequency <- lm(
  communication_apprehension ~ ride_frequency + public_frequency,
  data = data_all
)

summary(model_frequency)
```

**Each one-category increase in ride-share frequency was associated with an average 0.23-point decrease in communication apprehension, accounting for public-transport frequency, *b* = −0.23, *SE* = 0.05, *t*(249) = −4.28, *p* \< .001.**

**Each one-category increase in public-transport frequency was associated with an estimated 0.05-point decrease in communication apprehension after accounting for ride-share frequency; however, this association was not statistically significant, *b* = −0.05, *SE* = 0.05, *t*(249) = −1.00, *p* = .318.**

#### Visualization

```{r}
plot_data <- data_all |>
  select(
    communication_apprehension,
    ride_share_monthly,
    public_transport_monthly
  ) |>
  pivot_longer(
    cols = c(ride_share_monthly, public_transport_monthly),
    names_to = "transport",
    values_to = "frequency"
  ) |>
  mutate(
    frequency = as.numeric(as.character(frequency))
  )

ggplot(
  plot_data,
  aes(
    x = frequency,
    y = communication_apprehension,
    color = transport
  )
) +
  geom_jitter(width = 0.1, height = 0, alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Monthly use frequency",
    y = "Communication apprehension",
    color = "Transportation mode"
  ) +
  theme_minimal()
```

```{r}
exclusive_plot <- data_all |>
  filter(
    transport_group %in% c(
      "Ride share only",
      "Public transport only"
    )
  ) |>
  mutate(
    frequency = if_else(
      transport_group == "Ride share only",
      as.numeric(as.character(ride_share_monthly)),
      as.numeric(as.character(public_transport_monthly))
    )
  )

ggplot(
  exclusive_plot,
  aes(
    x = frequency,
    y = communication_apprehension,
    color = transport_group
  )
) +
  geom_jitter(width = 0.1, height = 0, alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Monthly use frequency",
    y = "Communication apprehension",
    color = "Transportation mode"
  ) +
  theme_minimal()
```

```{r}
exclusive_plot <- data_all |>
  filter(
    transport_group %in% c(
      "Ride share only",
      "Public transport only"
    )
  ) |>
  mutate(
    frequency = if_else(
      transport_group == "Ride share only",
      ride_frequency,
      public_frequency
    )
  )

all_plot <- data_all |>
  select(
    communication_apprehension,
    ride_frequency,
    public_frequency
  ) |>
  pivot_longer(
    cols = c(ride_frequency, public_frequency),
    values_to = "frequency"
  )

ggplot(
  exclusive_plot,
  aes(
    x = frequency,
    y = communication_apprehension,
    color = transport_group
  )
) +
  geom_jitter(
    width = 0.1,
    height = 0,
    alpha = 0.6
  ) +
  geom_smooth(
    method = "lm",
    se = FALSE
  ) +
  geom_smooth(
    data = all_plot,
    aes(
      x = frequency,
      y = communication_apprehension
    ),
    inherit.aes = FALSE,
    method = "lm",
    se = FALSE,
    color = "black"
  ) +
  labs(
    x = "Monthly use frequency",
    y = "Communication apprehension",
    color = "Transportation mode"
  ) +
  theme_minimal()
```
